---
title: "Hands-on Exercise 2"
author: "Lee Peck Khee"
date: 23 April 2023
date-modified: last-modified
edit: visual
---

## 2.1 Overview

In this chapter, you will be introduced to several ggplot2 extensions for creating more elegant and effective statistical graphics. By the end of this exercise, you will be able to:

-   control the placement of annotation on a graph by using functions provided in ggrepel package,

-   create professional publication quality figure by using functions provided in ggthemes and hrbrthemes packages,

-   plot composite figure by combining ggplot2 graphs by using patchwork package.

## 2.2 Getting Started

### 2.2.1 Install and loading the required libraries

In this exercise, beside tidyverse, four R packages will be used. They are:

-   ggrepel: an R package provides geoms for ggplot2 to repel overlapping text labels.

-   ggthemes: an R package provides some extra themes, geoms, and scales for 'ggplot2'.

-   hrbrthemes: an R package provides typography-centric themes and theme components for ggplot2.

-   patchwork: an R package for preparing composite figure created using ggplot2.

The code chunk below uses p_load() of pacman package to check if ggrepel, patchwork, ggthemes, hrbrthemes and tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse)
```

### 2.2.2 Importing the data

For the purpose of this exercise, a data file called Exam_data will be used. It consists of year end examination grades of a cohort of primary 3 students from a local school. It is in csv file format.

The code chunk below imports exam_data.csv into R environment by using read_csv() function of readr package. readr is one of the tidyverse package.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

There are a total of seven attributes in the exam_data tibble data frame. Four of them are categorical data type and the other three are in continuous data type.

-   The categorical attributes are: ID, CLASS, GENDER and RACE.

-   The continuous attributes are: MATHS, ENGLISH and SCIENCE.

## 2.3 Beyond ggplot 2 Annotation: ggrepel

One of the challenge in plotting statistical graph is annotation, especially with large number of data points.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

ggrepel is an extension of ggplot2 package which provides geoms for ggplot2 to repel overlapping text as in our examples on the right.

We simply replace geom_text() by geom_text_repel() and geom_label() by geom_label_repel.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## 2.4 Beyond ggplot2 Themes

ggplot2 comes with eight built-in themes, they are: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void().

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### 2.4.1 Working with ggtheme package

ggthemes provides 'ggplot2' themes that replicate the look of plots by Edward Tufte, Stephen Few, Fivethirtyeight, The Economist, 'Stata', 'Excel', and The Wall Street Journal, among others. In the example below, *The Economist* theme is used.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### 2.4.2 Working with hrbthemes package

hrbrthemes package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

The second goal centers around productivity for a production workflow. In fact, this "production workflow" is the context for where the elements of hrbrthemes should be used.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

## 2.5 Beyond Single Graph

It is not unusual that multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graphs. In this section, we create composite plot by combining multiple graphs. First, we create three statistical graphics by using the code chunk below.

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")

p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### 2.5.1 Creating Composite Graphics: pathwork methods

There are several ggplot2 extension's functions support the needs to prepare composite figure by combining several graphs such as grid.arrange() of gridExtra package and plot_grid() of cowplot package. In this section, we will use a ggplot2 extension called patchwork which is specially designed for combining separate ggplot2 graphs into a single figure.

Patchwork package has a very simple syntax where we can create layouts super easily. Here's the general syntax that combines:

-   Two-Column Layout using the Plus Sign +.

-   Parenthesis () to create a subplot group.

-   Two-Row Layout using the Division Sign /

### 2.5.2 Combining two ggplot2 graphs

Figure in the tabset below shows a composite of two histograms created using patchwork.

```{r}
p1 + p2
```

### 2.5.3 Combining three ggplot2 graphs

We can plot more complex composite by using appropriate operators. For example, the composite figure below is plotted by using:

-   "\|" operator to stack two ggplot2 graphs,

-   "/" operator to place the plots beside each other,

-   "()" operator the define the sequence of the plotting.

```{r}
(p1 / p2) | p3
```

### 2.5.4 Creating a composite figure with tag

In order to identify subplots in text, patchwork also provides auto-tagging capabilities as shown in the figure below.

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### 2.5.5 Creating figure with insert

Beside providing functions to place plots next to each other based on the provided layout. With inset_element() of patchwork, we can place one or several plots or graphic elements freely on top or below another plot.

```{r}
p12 <- p1|p2
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

### 2.5.6 Creating a composite figure by using patchwork and ggtheme

Figure below is created by combining patchwork and theme_economist() of ggthemes package discussed earlier.

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
