---
title: "In-class Exercise 4"
author: "Lee Peck Khee"
date: 6 May 2023
date-modified: last-modified
edit: visual
execute: 
  echo: true
  eval: true
  warning: false
---

```{r}
pacman:: p_load(rstatix, gt, patchwork, tidyverse)
```

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

### Visualizing Normal Distribution

-   A Q-Q plot, short for "quantile-quantile" plot, is used to assess whether or not a set of data potentially came from some theoretical distribution. In most cases, this type of plot is used to determine whether or not a data set follows a normal distribution.

-   If the data is normally distributied, the points in a Q-Q plot will lie on a straight diagonal line. Conversely, if the points deviate significantly from the straight diagonal line, then it's less likely that the data is normally distributed.

```{r}
ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```

In the section below, we present the information in via panel tabset to show two main tabs: "The plot" and "The code chunk".

::: panel-tabset
## The plot

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```

## The code chunk

```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```
:::

::: callout-note
We can see that the points deviate significantly from the straight diagonal line. This is a clear indication that the set of data is not normally distributed.
:::

### Combining statistical graph and analysis table

In the section below, we combine the statistical graph and analysis table into a single view for easy visualization.

::: panel-tabset
## The plot

```{r}
#| echo: false
qq <- ggplot(exam_data,
             aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png')
gtsave(sw_t,tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png
```

## The code chunk

```{r}
#| eval: false
qq <- ggplot(exam_data,
             aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png')
gtsave(sw_t,tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png
```
:::
